---
title: "ProjectP2"
author: "Me"
date: "23/10/2021"
output: pdf_document
---
## Load the ToothGrowth data and perform some basic exploratory data analyses


```{r}
#Load required packages
library(dplyr, warn.conflicts = F)
library(ggplot2)
library(ggthemes)
#Load data and convert to tbl format
ToothG <- tbl_df(ToothGrowth)
```

We take a look at the structure of our dataset and summarize the variables it contains

```{r}
#Structure of the dataframe
ToothG %>% str()
```

## Provide a basic summary of the data.

```{r}
#Summary
ToothG %>% summary()
sd(ToothG$len)
sd
```

```{r}
#Unique values in the dose vector
ToothG %>% select(dose) %>% unique()
```

The numeric variable _dose_ contains only 3 unique values: 0.5, 1, 2. We can conveniently convert it to a factor variable with three levels

```{r}
#Convert to factor
ToothG <- ToothG %>% mutate(dose = as.factor(dose))
```

####Plots

```{r, fig.height=3}
ToothG %>%
ggplot(aes(x=dose, y=len, fill = supp)) +
geom_boxplot() +
facet_grid(. ~ supp) +
scale_fill_brewer(palette = "Set1") +
theme_bw() +
ggtitle("Teeth Length vs Dose level \nby Supplement type") +
labs(x="dose(mg)", y= "teeth length ") +
guides(fill=guide_legend(title="Supplement type"))
```

This multipanel plot emphasizes the relationship between teeth length and dose level for each supplement type. It appears to be a positive relationship for both supplement types. In other words, as the amount of supplement increases, so does teeth length.

```{r, fig.height=3}
ToothG %>%
ggplot(aes(x = supp, y = len)) +
geom_boxplot(aes(fill = supp)) +
facet_wrap(~ dose) +
scale_fill_brewer(palette = "Set1") +
theme_bw() +
ggtitle("Teeth Length vs Supplement type \nby Dose level ") +
labs(x="supplement type", y= "teeth length ") +
guides(fill=guide_legend(title="Supplement type"))
```

This second plot shows the relationship between supplement type and teeth length emphasizing direct comparison between supplement types.

```{r}
ToothG  %>% filter(dose == 2)  %>% group_by(supp)   %>%  summarise(avg.length = mean(len))
```

We can see, at dosage level 2, VC appears to be slightly more effective than OJ, with an average teeth length of 26.14 vs 26.06

## Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)

**Testing by dose levels**

* #####Test A, dose = 0.5 and dose = 1

```{r}
#Exract the len and dose vectors from the df ToothGrowth
len_a <- ToothG %>% filter(dose %in% c(0.5,1)) %>% select(len) %>% unlist()
dose_a <- ToothG %>% filter(dose %in% c(0.5,1)) %>% select(dose) %>% unlist()
#Test
(Test.a <- t.test(len_a~dose_a, paired = FALSE))
```

* #####Test B, dose = 0.5 and dose = 2

```{r}
#Exract the len and dose vectors from the df ToothGrowth
len_b <- ToothG %>% filter(dose %in% c(0.5,2)) %>% select(len) %>% unlist()
dose_b <- ToothG %>% filter(dose %in% c(0.5, 2)) %>% select(dose) %>% unlist()
#Test
(Test.b <- t.test(len_b~dose_b, paired = FALSE))
```

* #####Test C, dose = 1 and dose = 2

```{r}
#Exract the len and dose vectors from the df ToothGrowth
len_c <- ToothG %>% filter(dose %in% c(1,2)) %>% select(len) %>% unlist()
dose_c <- ToothG %>% filter(dose %in% c(1,2)) %>% select(dose) %>% unlist()
#Test c
(Test.c <- t.test(len_c~dose_c, paired = FALSE))
```

We went through all possible combinations of levels from the factor variable dose and in all cases the p-value is lower than the default signficance level 0.05. Thus, we reject Ho. In other words there appears to be a positive relationship between dose level and teeth length

**Testing by Supplement**

```{r}
#Exract the len and supp vectors from the df ToothGrowth
len <- ToothG %>% select(len) %>% unlist()
supp <- ToothG %>% select(supp) %>% unlist()
#Test
t.test(len~supp, paired=F)
```

We can see that the p-value of the test is 0.06. Since the p-value is greater than 0.05 and the confidence
interval of the test contains zero, we can reject the null hypothesis and say that supplement types don't seem
to have any impact on teeth growth. In other words, there's no significant statistical difference between them

## State your conclusions and the assumptions needed for your conclusions.

Before using any statistical test we should always make sure that some conditions are met. In our case, t-tests, we should check for:

* Independence: there must be random sampling/assignment

* Normality: the population distribution must be normal or quasi-normal 

In this part of the project we can infer to the following points.

**Statistically, there is a significant difference between teeth length and dose levels across both delivery methods.Basically it means that as the dose increases so does teeth length.** 

**However,statistically there doesn't seem to be a significant difference between delivery methods. With OJ being apparently more effective at dose levels 0.5 and 1, and VC slightly more effective at dose level 2**